Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-13 paper/ijcai16.tex
paper: update methods section.

\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
capital letters $X$ for matrices, and calligraphic capital letters $\calX$ for
sets. The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$. We
use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.

\paragraph{Setting.} We assume to have a multi-attribute feature space $\calX$
of configurations $\vx = (x_1, \ldots, x_n)$ over $n$ features. For the time
being we will focus on binary features only, i.e. $x_z\in\{0,1\}$, to keep
matters simple; this restriction will be partially lifted later on.
\stefano{revisit the previous sentence once cost vars are in place}
We further assume the set of {\em feasible} configurations can be expressed as
a conjunction of linear constraints. This is not a severe restriction:
arbitrary propositional constraints can be trivially encoded in this setting,
e.g. the Boolean disjunction of two binary variables $x_1 \lor x_2$ can be
rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations will be
indicated as $\calX_\text{feasible}$.

In adherence to previous work\stefano{add reference}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we will
require all weights to be {\em non-negative} and {\em bounded}: the individual
weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$. These two requirements allow the translation of our core
optimization problem into a mixed-integer linear problem, as done in the next
section. \stefano{sentence is a bit bogus}

@@

\subsection{Problem Formulation}

@@

A straightforward encoding of the above desiderata leads to the following {\em
non-linear} optimization problem:
%
{\footnotesize
\begin{align}
    \max_{M, W, X}
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge M \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i} \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
}
%
Let us illustrate the above piece by piece. The objective is composed of four
parts: we want to maximize the shared margin $M$ (first part) and minimize the
total sum of the ranking errors incurred by each weight vector $\vw^{i}$
(second part), while at the same time promoting sparsity of the individual
weight vectors (third part) and quality of the configurations $\vx^{i}$ (last
part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
influence of the various components.

@@

Substituting the above MILP constraints into the original optimization problem,
we obtain:
{\footnotesize
\begin{align}
    \max_{M, W, X}
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M
    \\
        & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \nonumber
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \nonumber
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i}_h \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}

>>>>

\stefano{Paolo, feel free to improve the text and especially the citations
as you see fit.}

\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors, and
calligraphic capital letters $\calX$ for sets. We frequently abbreviate the set
$\{ x^i \}_{i=1}^n$ as $\{ x^i \}$ whenever the range of the index $i$ is clear
from the context, and use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$.

\paragraph{Setting.} We assume to have a multi-attribute feature space $\calX$
of configurations $\vx = (x_1, \ldots, x_n)$ over $n$ features. For the time
being we will focus on binary features only, i.e. $x_z\in\{0,1\}$, to keep
matters simple. This is a common choice for preference elicitation methods,
e.g.~\cite{guo2010real,viappiani2010optimal}; discrete features are handled by
means of a simple one-hot encoding. Partial support for continuous features
will be discussed later on.
We further assume that the set of {\em feasible} configurations is expressed as
a conjunction of linear constraints. This is not a severe restriction:
arbitrary propositional constraints can be trivially encoded in this setting,
e.g. the ``Boolean disjunction'' of two binary variables $x_1 \lor x_2$ can be
rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations will be
indicated as $\calX_\text{feasible}$.

In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we will
require all weights to be {\em non-negative} and {\em bounded}: the individual
weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$. Both requirements are quite natural, and enable the
translation of our core optimization problem into a mixed-integer linear
problem (as done in Section~\ref{sec:formulation}).

@@

\subsection{Problem Formulation}
\label{sec:formulation}

@@

A straightforward encoding of the above desiderata leads to the following {\em
non-linear} optimization problem over the variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$:
%
{\footnotesize
\begin{align}
    \max
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \label{eq:wyconstr}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge M \label{eq:wxconstr}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i} \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
}
%
Let us illustrate the above piece by piece. The objective is composed of four
parts: we want to maximize the shared margin $M$ (first part) and minimize the
total sum of the ranking errors incurred by each weight vector $\vw^{i}$
(second part), while at the same time promoting sparsity of the individual
weight vectors (third part) and quality of the configurations $\vx^{i}$ (last
part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
influence of the various components.

@@

Substituting the above MILP constraints into the original optimization problem,
we obtain:
{\footnotesize
\begin{align}
    \max
        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z
        \nonumber
    \\
    \text{s.t.}
        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
    \\
        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \nonumber
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M
    \\
        & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
    \\
        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
    \\
        & \;\; \forall \; i, j \in [k], i \neq j \nonumber
    \\
        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
    \\
        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \nonumber
    \\
        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \nonumber
    \\
        & \;\; \forall \; i \in [k] \;.\; \veps^{i}_h \ge 0 \nonumber
    \\
        & \;\; M \ge 0 \nonumber
\end{align}
